---
title: "covidHubUtils Multiple Models Exploration"
author: "Li Shandross"
date: "5/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(covidHubUtils)
library(tidyverse)
theme_set(theme_bw())
```


## Plot multiple models

Some additional arguments in `plot_forecasts()` are helpful for creating a reasonable plot if `forecast_data` has multiple locations, forecast dates or models.

The following code looks at three models' forecasts of incident deaths at one time point for one location. Note the use of the `fill_by_model` option which allows colors to vary by model and the `facet` command which is passed to ggplot.

```{r message=FALSE}
fdat <- load_latest_forecasts(models = c("Karlen-pypm", "UMass-MechBayes", "CU-select"),
                              last_forecast_date = "2021-03-08",
                              source = "zoltar",
                              forecast_date_window_size = 6,
                              locations = "US",
                              types = c("quantile", "point"), 
                              verbose = FALSE,
                              targets = paste(1:4, "wk ahead inc death"))

p <- plot_forecasts(fdat, 
                    target_variable = "inc death", 
                    truth_source = "JHU",
                    intervals = c(.5, .95), 
                    facet = .~model,
                    fill_by_model = TRUE, 
                    plot=FALSE) 

p +
  scale_x_date(name=NULL, date_breaks = "1 months", date_labels = "%b") +
  theme(axis.ticks.length.x = unit(0.5, "cm"),
    axis.text.x = element_text(vjust = 7, hjust = -0.2))
```


## Editing

```{r message=FALSE}
fdat <- 
  load_latest_forecasts(models = c("Karlen-pypm", "UMass-MechBayes", "CU-select"),
                              last_forecast_date = "2021-03-02",
                              source = "zoltar",
                              forecast_date_window_size = 6,
                              locations = "US",
                              types = c("quantile", "point"), 
                              verbose = FALSE,
                              targets = paste(1:4, "wk ahead inc death"))

p <- plot_forecasts(fdat, 
                    target_variable = "inc death", 
                    truth_source = "JHU",
                    intervals = c(.5, .95), 
                    facet = .~model,
                    fill_by_model = TRUE, 
                    plot=FALSE) 

p +
  scale_x_date(name=NULL, date_breaks = "1 months", date_labels = "%b") +
  theme(axis.ticks.length.x = unit(0.5, "cm"),
    axis.text.x = element_text(vjust = 7, hjust = -0.2))
```

```{r}
load_latest_forecasts(models = c("Karlen-pypm", "UMass-MechBayes", "CU-select"),
                              last_forecast_date = "2021-03-02",
                              source = "zoltar",
                              forecast_date_window_size = 6,
                              locations = "US",
                              types = c("quantile", "point"), 
                              verbose = FALSE,
                              targets = paste(1:4, "wk ahead inc death"))

###

#Understand the forecast format for models on the covid-19 forecast hub and write a function to construct a data frame with specific locations/date range.

my_function <- function(models, locations, targets){
  
}

```

```{r}
load_forecast_time_period <- function ( #function name change
  models = NULL,
  start_forecast_date_range, #
  end_forecast_date_range, #
  # removed forecast_date_window_size
  locations = NULL,
  types = NULL,
  targets = NULL,
  source = "local_hub_repo",
  hub_repo_path,
  as_of = NULL,
  hub = c("US", "ECDC"),
  verbose = TRUE) {
  
  # validate source
  source <- match.arg(source, choices = c("local_hub_repo", "zoltar"))
  
  # check date format and generate dates of forecasts to load
  forecast_dates <- tryCatch({
   as.character(as.Date(start_forecast_date_range)) 
    as.character(as.Date(end_forecast_date_range))
    as.character(seq(as.Date(start_forecast_date_range), as.Date(end_forecast_date_range), by = "7 days"))
    }, error = function(err){
    }, error = function(err){
      stop("Error in load_latest_forecasts: Please provide a valid date object or
           string in format YYYY-MM-DD in start_forecast_date_range or end_forecast_date_range")
      }
    )

  if (source == "local_hub_repo") {
    # validate hub repo path
    if (missing(hub_repo_path) | !dir.exists(hub_repo_path)) {
      stop("Error in load_latest_forecasts: Please provide a vaid path to hub repo.")
    } 
    
    # path to data-processed folder in hub repo
    data_processed <- file.path(hub_repo_path, "data-processed/")
    
    forecasts <- load_latest_forecasts_repo(file_path = data_processed, 
                                            models = models, 
                                            forecast_dates = forecast_dates, 
                                            locations = locations, 
                                            types = types, 
                                            targets = targets, 
                                            hub = hub, 
                                            verbose = verbose)
  } else {
    forecasts <- load_latest_forecasts_zoltar(models = models, 
                                              forecast_dates = forecast_dates,
                                              locations = locations, 
                                              types = types,
                                              targets = targets,
                                              as_of = as_of,
                                              hub = hub,
                                              verbose = verbose)
  }
  
  return(forecasts)
}


load_forecast_time_period(
  models = "Karlen-pypm",
  source = "zoltar",
  start_forecast_date_range = "2021-03-02", 
  end_forecast_date_range = "2021-03-09",
  locations = "US",
  types = "point",
  targets = "1 wk ahead inc death",
  verbose = FALSE)

##
 load_latest_forecasts(models = c("Karlen-pypm"),
                              last_forecast_date = "2021-03-02",
                              source = "zoltar",
                              forecast_date_window_size = 6,
                              locations = "US",
                              types = c("point"), 
                              verbose = FALSE,
                              targets = paste(1, "wk ahead inc death"))
 

```


```{r}
tryCatch({
   as.character(as.Date(start_fdr)) 
    as.character(as.Date(end_fdr))
    as.character(seq(as.Date(start_fdr), as.Date(end_fdr), by = "7 days"))
    }, error = function(err){
      stop("Error in load_latest_forecasts: Please provide a valid date object or
           string in format YYYY-MM-DD in start_forecast_date_range or end_forecast_date_range")
      }
    )


tryCatch({
    as.character(as.Date("2021-03-02") +
                   seq(from = -6, to = 0, by = 1))
    }, error = function(err){
      stop("Error in load_latest_forecasts: Please provide a valid date object or
           string in format YYYY-MM-DD in latest_forecast_date.")
      }
    )
```

